---
title: "Analysis of Movie Ratings Trends Over Time"
author: "Prasham"
date: "03/17/2025"
---
# Loading Libraries and Data:
```{r message=FALSE}
library(readr)

previous <- read_csv('fandango_score_comparison.csv')
after <- read_csv('movie_ratings_16_17.csv')

head(previous)
```

# Previewing the after Dataset:
```{r}
head(after)
```
# Data Selection and Preview for Fandango Movie Ratings Analysis:
I selected specific columns from the previous and after datasets to focus on important information such as movie names, ratings, and votes. This helped simplify the analysis by narrowing down the data. Then, I previewed the first few rows to ensure the selection was correct and the data was ready for further analysis.

```{r message=FALSE}
library(dplyr)
fandango_previous <- previous %>% 
  select(FILM, Fandango_Stars, Fandango_Ratingvalue, 
         Fandango_votes, Fandango_Difference)

fandango_after <- after %>% 
  select(movie, year, fandango)

head(fandango_previous)
```
# Previewing the fandango_after Dataset:
```{r}
head(fandango_after)
```
# Random Sampling of Movies from the Fandango Dataset:
I used the set.seed() function to ensure reproducibility and then randomly selected 10 movies from the fandango_after dataset using sample_n(). This step helps to create a smaller, manageable subset of data for quick checks or for further detailed analysis without working with the entire dataset.

```{r}
set.seed(1)
sample_n(fandango_after, size = 10)
```
# Adding Rotten Tomato Review Counts to Random Sampled Movies:
I first randomly selected 10 movies from the fandango_after dataset using sample_n(), ensuring reproducibility with set.seed(). Then, I created a new dataset containing the Rotten Tomato review counts for each of the selected movies. Finally, I combined the two datasets using bind_cols() to add the review counts to the randomly sampled movie data, making it ready for analysis.
```{r}
set.seed(1)
sampled <- sample_n(fandango_after, size = 10)
reviews <- tibble(reviews = c(13569, 74904, 24293, 4141, 30183, 48952, 14328, 59359, 54765, 82222))
bind_cols(sampled, reviews)
```

# Counting Movies with Fewer Than 30 Votes:
```{r}
sum(fandango_previous$Fandango_votes < 30)
```

# Previewing the First 10 Movie Titles:
```{r}
head(fandango_previous$FILM, n = 10)
```

# Identifying Unique Years in the Fandango After Dataset:
```{r}
unique(fandango_after$year)
```

# Extracting Year from Movie Titles:
```{r}
library(stringr)
fandango_previous <- fandango_previous %>% 
  mutate(year = str_sub(FILM, -5, -2))
```

# Grouping Movies by Year and Counting Frequency:
```{r}
fandango_previous %>% 
  group_by(year) %>% 
  summarize(Freq = n())
```

# Creating a Frequency Table for Movie Years
```{r}
table(fandango_previous$year)
```

# Filtering Data for 2015 Movies:
```{r}
fandango_2015 <- fandango_previous %>% 
  filter(year == 2015)
table(fandango_2015$year)
```

# Previewing the First 6 Movies in the Fandango After Dataset:
```{r}
head(fandango_after)
```

# Creating a Frequency Table for Movie Years in the Fandango After Dataset:
```{r}
table(fandango_after$year)
```

# Filtering Data for 2016 Movies:
```{r}
fandango_2016 <- fandango_after %>% 
  filter(year == 2016)
table(fandango_2016$year)
```

# Creating a Density Plot to Compare Fandango Ratings in 2015 vs 2016:
I created a density plot to compare the distribution shapes of Fandango ratings for 2015 and 2016. The plot overlays the density of ratings for both years, allowing for a clear visual comparison of the distribution. The 2015 ratings are represented by the default color, while the 2016 ratings are shown in blue.
```{r}
library(ggplot2)
ggplot(data = fandango_2015, 
               aes(x = Fandango_Stars)) +
  geom_density() +
  geom_density(data = fandango_2016, 
               aes(x = fandango), color = "blue") +
  labs(title = "Comparing distribution shapes for Fandango's ratings\n(2015 vs 2016)",
       x = "Stars",
       y = "Density") +
  scale_x_continuous(breaks = seq(0, 5, by = 0.5), 
                     limits = c(0, 5))
```

# Calculating the Percentage of Each Fandango Rating for 2015 Movies:
I calculated the percentage of each Fandango rating (from 3.0 to 5.0) for the movies released in 2015. This helps to understand the distribution of ratings and how frequently each rating occurred among the popular movies in that year.
```{r}
fandango_2015 %>% 
  group_by(Fandango_Stars) %>% 
  summarize(Percentage = n() / nrow(fandango_2015) * 100)
```

# Calculating the Percentage of Each Fandango Rating for 2016 Movies:
I calculated the percentage of each Fandango rating (from 3.0 to 5.0) for the movies released in 2016. This helps to understand the distribution of ratings and how frequently each rating occurred among the popular movies in that year.
```{r}
fandango_2016 %>% 
  group_by(fandango) %>% 
  summarize(Percentage = n() / nrow(fandango_2016) * 100)
```

# Calculating Summary Statistics for 2015 and 2016 Fandango Ratings:
I first defined a custom mode function to calculate the most frequent rating value for the datasets. Then, I calculated the mean, median, and mode for Fandango ratings in 2015 and 2016. I combined the summary statistics for both years into one dataset and reshaped it into a format suitable for visualization with ggplot.
```{r}
library(tidyr)

mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

summary_2015 <- fandango_2015 %>% 
  summarize(year = "2015",
    mean = mean(Fandango_Stars),
    median = median(Fandango_Stars),
    mode = mode(Fandango_Stars))

summary_2016 <- fandango_2016 %>% 
  summarize(year = "2016",
            mean = mean(fandango),
            median = median(fandango),
            mode = mode(fandango))

summary_df <- bind_rows(summary_2015, summary_2016)

summary_df <- summary_df %>% 
  gather(key = "statistic", value = "value", - year)

summary_df
```

# Visualizing Summary Statistics for 2015 and 2016 Fandango Ratings:
I created a bar plot to compare the mean, median, and mode of Fandango ratings for 2015 and 2016. The bars represent each statistic (mean, median, and mode), with different colors indicating the two years. The chart provides a visual comparison of the distribution of ratings across the two years.
```{r}
ggplot(data = summary_df, aes(x = statistic, y = value, fill = year)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparing summary statistics: 2015 vs 2016",
       x = "",
       y = "Stars")
```

# Calculating the Change in Mean Fandango Ratings from 2015 to 2016:
I filtered the dataset to focus on the mean ratings for 2015 and 2016 and calculated the percentage change in the mean rating between the two years. The result shows a slight increase in the mean rating from 2015 to 2016.
```{r}
means <- summary_df %>% 
  filter(statistic == "mean")

means %>% 
  summarize(change = (value[1] - value[2]) / value[1])
```
# Conclusion
The analysis revealed a small but noticeable difference between the Fandango ratings for popular movies in 2015 and those in 2016. On average, popular movies released in 2016 received lower ratings on Fandango compared to their 2015 counterparts. This shift can be attributed to Fandango addressing the bias in its rating system following Hickey's analysis.
